Interpretable Machine Learning

### 2018-08-17 talk-40min.Rmd
improves talk

```{r, echo=FALSE, out.height = '100%', out.width=NULL}
knitr::include_graphics("images/wow.jpg")
```

@@

```{r, echo=FALSE, out.height = '100%', out.width=NULL}
knitr::include_graphics("images/magic.jpg")
```

## Step 1: Find some data
Kaggle conducted an industry-wide survey of data scientists. 
https://www.kaggle.com/kaggle/kaggle-survey-2017

@@

## Step 2: Throw ML on your data
```{r learn}
library('mlr')
set.seed(42)
task = makeRegrTask(data = survey.dat, target = 'CompensationAmount')
lrn = makeLearner('regr.randomForest')
mod = train(lrn, task)
```

```{r, echo=FALSE}
knitr::include_graphics("images/comp-dog.gif")
```

## Step 3: Profit. {.center}

```{r, echo=FALSE, out.height=NULL, out.width='100%'}
knitr::include_graphics("images/done-here.gif")
```

@@

"What problem?""

```{r, echo=FALSE, out.height='50%'}
knitr::include_graphics("images/Hide-the-pain-harold-phone.jpg")
```

@@

```{r, echo=FALSE, out.height='30%'}
knitr::include_graphics("images/age.jpeg")
```

# Partial Dependence Plots

```{r}
library("iml")
X = survey.dat[-which(names(survey.dat) == "CompensationAmount")]
predictor = Predictor$new(mod, data = X, y = getTaskTargets(task))
pd = Partial$new(predictor, feature ='Age')
pd$plot() 
```

@@

@@

# Permutation feature importance

@@

## Gender?!
```{r, echo=FALSE, out.width='80%'}
knitr::include_graphics("images/big-mistake.png")
```

@@

```{r, echo=FALSE}
knitr::include_graphics("images/hidden-pain-bias.jpg")
```

```{r, echo=FALSE, out.width='40%'}
knitr::include_graphics("images/angry.gif")
```

@@

@@

# Shapley

@@

```{r, echo=FALSE, fig.align='center'}
knitr::include_graphics("images/black-box.png")
```

- Opaque decision making by machine learning algorithms
- Bias and mistakes in the data
- Trust as a user: Should I start a therapy with sever side effects, because the machine said so?
- Debugging as a practitioner: Why did the algorithm miss-classify sample X? Did it learn generalizable features?

@@

Interpretability is the degree to which a human can understand the cause of a decision. 
But we don't really have a good way to measure that really. Not as easy as benchmarking ML algorithms.

@@

<font size="2">Miller, Tim. 2017. “Explanation in Artificial Intelligence: Insights from the Social Sciences.” arXiv Preprint arXiv:1706.07269.</font>

@@

We need interpretability when the loss function does not cover all constraints.

- Decisions about humans.
- Critical applications that decide about life and death.
- Newly developed systems with unknown consequences.
- Models using proxies instead of causal inputs.
- Debugging the models.
- Increasing trust.

### Examples where we need interpretability

- Credit scores
- Autonomous vehicles
- Medical recommendations
- Automated diagnoses

@@

## Example: Digit recognition

![](images/mnist.jpeg)

@@

@@

- Decision trees
- Decision rules
- Generalized regression models
- ...

@@

- RandomForestExplainer
- Visualizing activations of neural networks
-

TODO: add images from papers

@@

- Partial dependence plots
- LIME
- ...

@@

- Prototypes: What is the prototype of a certain class
- Criticisms: Which data points are unusual for a certain class. 
- Counterfactuals: What do I need to change for an instance to change the prediction?
- Anchors: What features do I have to anchor to keep prediction the same?

TODO: Cite

## Outcomes of interpretability method:

- Model
- Data point
- Interpretable unit (like a weight)
- Summary statistic for a feature

>>>>

## {.center data-background=images/wow.jpg data-background-size=contain}

@@

## {data-background=images/magic.jpg data-background-size=contain}

# Step 1: Find some data

## Step 1: Find some data

Kaggle conducted an industry-wide survey of data scientists. 
https://www.kaggle.com/kaggle/kaggle-survey-2017

@@

# Step 2: Throw ML on your data

```{r learn}
library('mlr')
set.seed(42)
task = makeRegrTask(data = survey.dat, target = 'CompensationAmount')
lrn = makeLearner('regr.randomForest')
mod = train(lrn, task)
```

## {.center data-background=images/comp-dog.gif data-background-size=contain}

<div class="notes">
- Random Forest
- Target: Compensation
- All features that you found in the data
- Train, test, ship
</div>

# Step 3: Profit. {.center}

## {data-background=images/done-here.gif data-background-size=contain}
```{r, echo=FALSE, out.width='80%', include = FALSE}
knitr::include_graphics("images/done-here.gif" )
```

@@

##  {.emphasizedabit .center data-background=images/Hide-the-pain-harold-phone.jpg data-background-size=cover}

"What problem?"

```{r, echo=FALSE, out.height='50%', include=FALSE}
knitr::include_graphics("images/Hide-the-pain-harold-phone.jpg")
```

@@

```{r, echo=FALSE, fig.width=16, include=FALSE}
knitr::include_graphics("images/age.jpeg")
```

##  {.emphasizedabit .center data-background=images/black-box.gif data-background-size=cover}

<div class="white">
Looking inside the black box
</div>

# How do the features influence my predictions?

## Partial Dependence Plot

```{r out.width="80%"}
library("iml")
X = survey.dat[-which(names(survey.dat) == "CompensationAmount")]
predictor = Predictor$new(mod, data = X, y = getTaskTargets(task))
pd = Partial$new(predictor, feature ='Age')
pd$plot() 
```

@@

<div class="notes">
- PDP answers: How does the prediction change when a feature input changes?
- Average of ICE curves
</div>

@@

# What are the most important features?

## Permutation feature importance

@@

## Gender?! {.emphasizedabit .center data-background=images/big-mistake.png data-background-size=contain}

```{r, echo=FALSE, out.width='80%', include=FALSE}
knitr::include_graphics("images/big-mistake.png")
```

# What's the influence of gender on the prediction?

@@

## {.center data-background=images/hidden-pain-bias.jpg data-background-size=contain}

```{r, echo=FALSE, include = FALSE}
knitr::include_graphics("images/hidden-pain-bias.jpg")
```

## {.center data-background=images/angry.gif data-background-size=contain}

```{r, echo=FALSE, out.width='40%', include = FALSE}
knitr::include_graphics("images/angry.gif")
```

# Explaining individual predictions

@@

## {.center data-background=images/lime-fitting-1.png data-background-size=contain}

@@

# Shapley Value

TODO: find image for explaining shapley value

@@

```{r, echo=FALSE, fig.align='center'}
knitr::include_graphics("images/black-box.png")
```
<div class="notes">
- Opaque decision making by machine learning algorithms
- Bias and mistakes in the data
- Trust as a user: Should I start a therapy with sever side effects, because the machine said so?
- Debugging as a practitioner: Why did the algorithm miss-classify sample X? Did it learn generalizable features? 
</div>

@@

<div class="notes">
Interpretability is the degree to which a human can understand the cause of a decision. 
But we don't really have a good way to measure that really. Not as easy as benchmarking ML algorithms.

@@

<font size="2">Miller, Tim. 2017. “Explanation in Artificial Intelligence: Insights from the Social Sciences.” arXiv Preprint arXiv:1706.07269.</font>
</div>

@@

<div class="notes">
We need interpretability when the loss function does not cover all constraints.

- Decisions about humans.
- Critical applications that decide about life and death.
- Newly developed systems with unknown consequences.
- Models using proxies instead of causal inputs.
- Debugging the models.
- Increasing trust.
</div>

## Examples where we need interpretability

TODO: Insert images
- Credit scores
- Autonomous vehicles
- Medical recommendations
- Automated diagnoses

@@

## Example: Digit recognition {.center data-background=images/mnist.jpeg data-background-size=cover}

@@

TODO: replace with image

@@

<div class="notes">
- Decision trees
- Decision rules
- Generalized regression models
- ...
</div>

@@

<div class="notes">
- Visualizing activations of neural networks
- Random Forest specific feature importance
- xgboost explainer
- problem: new models, not comparable between models, missing flexibility
</div>

@@

<div class="notes">
- Partial dependence plots
- LIME
- ...
</div>

@@

<div class="notes">
- Prototypes: What is the prototype of a certain class
- Criticisms: Which data points are unusual for a certain class. 
- Counterfactuals: What do I need to change for an instance to change the prediction?
- Anchors: What features do I have to anchor to keep prediction the same?
</div>

### 2018-08-17 talk.Rmd
improves talk

## "The older the applicants, the higher the predicted salary, regardless of skills."
```{r, echo=FALSE, out.width='100%', fig.align='center'}
knitr::include_graphics("images/age.jpeg")
```

@@

```{r}
library("iml")
X = survey.dat[-which(names(survey.dat) == "CompensationAmount")]
predictor = Predictor$new(mod, data = X, y = getTaskTargets(task))
ice = Ice$new(predictor, feature ='Age')
ice$plot()
```

>>>>

## "The older the applicants, the higher the predicted salary, regardless of skills."
```{r, echo=FALSE, out.height='100%', fig.align='center'}
knitr::include_graphics("images/age.jpeg")
```

@@

```{r}
library("iml")
X = survey.dat[-which(names(survey.dat) == "CompensationAmount")]
predictor = Predictor$new(mod, data = X, y = getTaskTargets(task))
ice = Partial$new(predictor, feature ='Age')
ice$plot()
```
